---
title: "PA1_assignment"
author: "GMW657"
date: "February 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is a reproducible research assignment(project 1)

set working directory (change this to fit your needs)
```{r}
setwd("E:/coursera/RepData_PeerAssessment1")
```
load all packages used in this exploratory analysis

```{r}
library(knitr)
library(dplyr)
library(ggplot2)
opts_chunk$set(echo = TRUE)
```
##Loading and preprocessing the data
unzip file
```{r}
unzip(zipfile = "activity.zip")
```
load data
```{r}
data_row <- read.csv('activity.csv')

# remove NA in data
data <- data_row[ with (data_row, { !(is.na(steps)) } ), ]
```
print out first 20 rows

```{r}
head(data,20)
```
##mean total number of steps taken per day
```{r}
by_day <- group_by(data, date)
steps_by_day <- summarise(by_day, total = sum(steps))
steps_by_day
```
Calculate and report the mean and median of the total number of steps taken per day
```{r}
hist(steps_by_day$total, main="Histogram of total number of steps per day", 
     xlab="Total number of steps in a day")
summary(steps_by_day)
```
##The average daily activity pattern
 preprocessing data for plot
```{r}
 steps_by_interval <- aggregate(steps ~ interval, data, mean)
```
Create a time series plot 
```{r}
plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
     main="Average number of steps over all days", xlab="Interval", 
     ylab="Average number of steps")
```
 Find row with max of steps
```{r}
max_steps_row <- which.max(steps_by_interval$steps)
```
Find interval with this max
```{r}
steps_by_interval[max_steps_row, ]
```
##Imputing missing values
Calculate the total number of rows with NAs
```{r}
sum(is.na(data_row))
```
Devise a strategy for filling in all of the missing values in the dataset. 
```{r}
data_imputed <- data_row
for (i in 1:nrow(data_imputed)) {
  if (is.na(data_imputed$steps[i])) {
    interval_value <- data_imputed$interval[i]
    steps_value <- steps_by_interval[
      steps_by_interval$interval == interval_value,]
    data_imputed$steps[i] <- steps_value$steps
  }
}
```
calculate  total number of steps taken each day
```{r}
df_imputed_steps_by_day <- aggregate(steps ~ date, data_imputed, sum)
head(df_imputed_steps_by_day)
```
Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r}
hist(df_imputed_steps_by_day$steps, main="Histogram of total number of steps per day (imputed)", 
     xlab="Total number of steps in a day")
```
get mean and median of imputed data
```{r}
mean(df_imputed_steps_by_day$steps)

median(df_imputed_steps_by_day$steps)
```
Get mean and median of data without NA's
```{r}
mean(steps_by_day$total)
median(steps_by_day$total)
```
## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
data_imputed['type_of_day'] <- weekdays(as.Date(data_imputed$date))
data_imputed$type_of_day[data_imputed$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
data_imputed$type_of_day[data_imputed$type_of_day != "weekend"] <- "weekday"
```
convert type_of_day from character to factor
```{r}
data_imputed$type_of_day <- as.factor(data_imputed$type_of_day)
```
calculate average steps by interval across all days
```{r}
df_imputed_steps_by_interval <- aggregate(steps ~ interval + type_of_day, data_imputed, mean)
```
creat a plot
```{r}
qplot(interval, 
      steps, 
      data = df_imputed_steps_by_interval, 
      type = 'l', 
      geom=c("line"),
      xlab = "Interval", 
      ylab = "Number of steps", 
      main = "") +
  facet_wrap(~ type_of_day, ncol = 1)
```

